---
title: "London Analysis 1"
author: "London Wagner"
date: "11/8/2020"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggplot2)
```

## Abstract
The goal of this exploratory analysis is to examine the correlation
between the quantitative variables in the "water" data set. At this 
point, we have decided that we are going to be using the number of
violations as the response variable in our model, so I plan on checking 
the relationship between the quantitative predictors. 

## Analysis
There are 14 variables in this dataset, but for this analysis, we will be
focusing on the following ones:

- `population`: a double that represents the population a site caters to
- `num_facilities`: a double that represents the number of facilities at a site
- `num_site_visits`: a double that represents how many times the site has been
visited

```{r}
water <- read.csv("../data/Tidy_Data/tidy_water.csv")
water$region <- as.factor(water$region)
head(water)
```

# Number of Facilities and Number of Site Visits
```{r}
ggplot(data = water)+
  geom_point(mapping = aes(x = num_facilities, y = num_site_visits))
```
Based on the initial scatterplot, it is difficult to tell if there is 
any sort of correlation. The majority of the observations appear to have
be centered around a relatively low number of site visits and low number
of facilities. There are observations with a low number of site visits
and a high number of facilities, but there are only around 6 of these 
out of the 100,000+ observations. There are also some points with a low
number of facilities and a high number of site visits. Overall, there 
is no clear pattern shown. 

```{r}
ggplot(data = water)+
  geom_point(mapping = aes(x = log(num_facilities + 1), y = log(num_site_visits + 1)))
```
After transforming the data, there is once again no clear pattern between the
two variables.

# Number of Facilities and Population
```{r}
ggplot(data = water)+
  geom_point(mapping = aes(x = num_facilities, y = population))
```
The initial scatterplot shows that there might be a slight positive 
correlation between population and number of facilities, but at
this point I am not sure. I'm going to transform the data to hopefully 
provide more insight.

```{r}
ggplot(data = water)+
  geom_point(mapping = aes(x = log(num_facilities + 1), y = log(population +1)))
```
After transforming the data, a positive correlation becomes more clear. As
the number of facilities increases, it appears that the population increases
as well. We saw earlier that the number of site visits did not have much of
an impact on the number of facilities, but we do not know if it has an impact
on population.

# Population and Number of Site Visits

```{r}
ggplot(data = water)+
  geom_point(mapping = aes(x = population, y = num_site_visits))
```
Once again, it is difficult to tell if there is any correlation between the 
two variables, so I will transform them in the same way I have previously to
get a better understanding of whether or not there is a potential relationship 
between the two.

```{r}
ggplot(data = water)+
  geom_point(mapping = aes(x = log(population + 1), y = log(num_site_visits+1)))
```
There is a slight positive association between the two transformed 
variables. As population increases, the general trend is that the
number of site visits increases. 

## Conclusion
The variable population appears to be at least slightly associated with
the number of site visits and the number of facilities. However, the 
number of site visits and the number of facilities to not appear to have
an association. If there is an association between population and our 
response, the number of violations, this could be coming from its 
correlation with the other two quantitative predictors.

