---
title: "London Analysis 1"
author: "London Wagner"
date: "11/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggplot2)
```

## Abstract
The goal of this exploratory analysis is to examine the "water" data set and 
determine whether the EPA region a site is located in (region) will have an
effect on the number of violations a site will have (num_violations).

## Analysis
There are 14 variables in this dataset, but for this analysis, we will be
focusing on the following ones:

- `region`: a factor that represents which EPA region the site is in (the
region determines which office is responsible for the execution of the EPA's
programs)
- `population`: a double that represents the population a site caters to
- `num_facilities`: a double that represents the number of facilities at a site
- `num_violations`: a double that represents the number of violations at a site
- `num_site_visits`: a double that represents how many times the site has been
visited

```{r}
water <- read.csv("../data/Tidy_Data/tidy_water.csv")
head(water)
```

# Converting Region into a factor and checking it
```{r}
water$region <- as.factor(water$region)
levels(water$region)
```

# Univariate Analysis
First, we'll create a bar chart for the number of facilities in each region
```{r}
water%>%
  ggplot(aes(x = region, fill = region))+
  geom_bar()+
  theme_bw()+
  ggtitle("Count per EPA Region")+
  theme(legend.position = "none")
```
There appear to be the greatest number of sites in region 5 at around 42,000.
The number of sites in region 7 and 8 are the lowest, at around 7,000. 

Next, we'll create a histogram for the distribution of the number of violations

```{r}
water%>%
  ggplot(aes(x = num_violations))+
  geom_histogram(bins = 30, fill = "light blue")+
  theme_bw()+
  ggtitle("Number of Violations")
```
The data is heavily skewed, so we will transform it with a log transformation.
We will also have to add 1, since there are 0 values.

```{r}
water%>%
  ggplot(aes(x = log(num_violations + 1)))+
  geom_histogram(bins = 30, fill = "light blue")+
  theme_bw()+
  ggtitle("Number of Violations")
```
Even though the distribution is still not normal, it is much closer to normal
than it was before. Additionally, because our sample size is so large, it's
safe to assume that normality will not be an issue for this data set. However,
because there are a large proportion of 0's in this data set, there could be
problems in statistical analysis.

# Computing summary statistics by group
```{r}
water%>%
  group_by(region)%>%
  summarise(
    count = n(),
    mean = mean(num_violations),
    sd = mean(num_violations)
  )
```

# Bivariate Analysis
```{r}
water%>%
  ggplot(aes(x = region, y = log(num_violations+1), fill = region))+
  geom_boxplot()+
  theme_bw()+
  theme(legend.position = "none")+
  ggtitle("Region vs. Number of Violations(log)")
```
There definitely appears to be variation in the number of violations depending
on the region the site is located in. We will need to examine this further. We
also need to take into account the number of facilities at each site. More
violations could occur in each region due to the fact that there tend to be
more facilities per site. For now, let's just examine the difference in
violations across regions with an ANOVA table. Not all the assumptions will 
be 100% valid for the ANOVA table, but we will use it for inital explorations.

```{r}
anova(lm(num_violations~region, data = water))
```
Because our p-value is less than any significant alpha, we can conclude that 
there is a difference in the mean number of violations per region.

I'm going to create a variable that determines the number of violations based
on the number of facilities at each site. To do this, I will be dividing the
number of violations by the number of facilities, therefore getting the number
of violations per facilities. 
```{r}
water%>%
  filter(num_facilities > 0)%>%
  mutate(violations_per_fac = num_violations/num_facilities)-> water
```

There appear to be 22 sites with no facilities. At this point, I'm not quite
sure why that is, but because these points make up such a small percentage of
the data set, we are going to assume these data points will not change our
analysis and leave them out.

```{r}
water%>%
  ggplot(aes(x = region, y = log(violations_per_fac+1), fill = region))+
  geom_boxplot()+
  theme_bw()+
  theme(legend.position = "none")+
  ggtitle("Region vs. Number of Violations(log)")
```

```{r}
anova(lm(violations_per_fac~region, data = water))
```

Because our p-value is less than any significant alpha, we can conclude that 
there is a difference in the mean number of violations per region, even when 
factoring in the number of facilities.

## Conclusion

There appears to be a difference in the mean number of violations based
on region, even when factoring in the number of facilities. However, we 
do not know if there are other variables within the data set that also 
contribute to the number of violations. This will need to be examined in
further analysis when we try to create a model for the data. 

