---
title: "Regression Models"
author: "Favian Liu"
date: "12/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(ggplot2)
library(shinythemes)
library(tidyverse)
library(broom)
library(arm)
library(leaps)
library(lmtest)
library(car)
library(sandwich)
water <- read_csv("../output/tidy_water.csv")
```

## Tidying the data further for analysis
```{r}
glimpse(water)
typeof(water)
water %>%
  mutate(region = as.factor(region)) %>%
         filter(source == "Ground water" | source == "Surface water") %>%
  mutate(source = as.factor(source)) %>%
  dplyr::select(region, primary_agency, population, num_facilities, num_violations, num_site_visits, source) -> water

summary(water)
```


## Running a Poisson Analysis
```{r}
glm.water <- glm(num_violations ~ population + num_facilities + num_site_visits + source + region, data = water, family = poisson(link = "log"))

glm.water %>%
  augment() -> water.resid

water.resid %>%
  ggplot(aes(x = .fitted, y = .resid)) +
  geom_point()

summary(glm.water)

cov.water <- vcovHC(glm.water, type="HC0")
std.err <- sqrt(diag(cov.water))
r.est <- cbind(Estimate= coef(glm.water), "Robust SE" = std.err,
               "Pr(>|z|)" = 2 * pnorm(abs(coef(glm.water)/std.err), lower.tail=FALSE),
               LL = coef(glm.water) - 1.96 * std.err,
               UL = coef(glm.water) + 1.96 * std.err)

r.est

```

The poisson model shows the estimate for the variables are all significant. However, based on the residual plot of the poisson analysis, this model is a bad fit. A quasipoisson model might be better.

## Running a Quasipoisson Analysis

```{r}
quasi.water <- glm(num_violations ~ population + num_facilities + num_site_visits + source + region, data = water, family = quasipoisson(link = "log"))

quasi.water %>%
  augment() -> quasi.resid
  
summary(quasi.water)
```
A quasipoisson model shows us the estimates of the same variables are no longer significant.

# Exploring the best model to use


## Running a stepwise by calculating "quasi-AIC"
Since R does not give the AIC value of quasipoisson models, I used this website https://cran.r-project.org/web/packages/bbmle/vignettes/quasi.pdf to try to determine the AIC values. But I was having trouble with the function.

```{r}
glmFVSR.water <- update(glm.water, . ~ . - population)
glmFVR.water <- update(glmFVSR.water, . ~ . - source)
glmFSR.water <- update(glmFVSR.water, . ~ . - num_site_visits)
glmVSR.water <- update(glmFVSR.water, . ~ . - num_facilities)
glmFVS.water <- update(glmFVSR.water, . ~ . - region)
glmFR.water <- update(glmFVR.water, . ~ . -num_site_visits)
glmVR.water <- update(glmVSR.water, . ~ . - source)
glmFV.water <- update(glmFVS.water, . ~ . - source)
glmRS.water <- update(glmFSR.water, . ~ . - num_facilities)
glmFS.water <- update(glmFVS.water, . ~ . - num_site_visits)
glmVS.water <- update(glmFVS.water, . ~ . - num_facilities)
glmR.water <- update(glmFR.water, . ~ . - num_facilities)
glmF.water <- update(glmFV.water, . ~ . - num_site_visits)
glmV.water <- update(glmFV.water, . ~ . - num_facilities)
glmS.water <- update(glmVS.water, . ~ . - num_site_visits)
glmFVSR.water <- update(glmFVSR.water, family=quasipoisson)
glmFVR.water <- update(glmFVR.water, family=quasipoisson)
glmFSR.water <- update(glmFSR.water, family=quasipoisson)
glmVSR.water <- update(glmVSR.water, family=quasipoisson)
glmFVS.water <- update(glmFVS.water, family=quasipoisson)
glmFR.water <- update(glmFR.water, family=quasipoisson)
glmVR.water <- update(glmVR.water, family=quasipoisson)
glmFV.water <- update(glmFV.water, family=quasipoisson)
glmRS.water <- update(glmRS.water, family=quasipoisson)
glmFS.water <- update(glmFS.water, family=quasipoisson)
glmVS.water <- update(glmVS.water, family=quasipoisson)
glmR.water <- update(glmR.water, family=quasipoisson)
glmF.water <- update(glmF.water, family=quasipoisson)
glmV.water <- update(glmV.water, family=quasipoisson)
glmS.water <- update(glmS.water, family=quasipoisson)

summary(glmFVSR.water)
summary(glmFVR.water)
summary(glmFSR.water)
summary(glmVSR.water)
summary(glmFVS.water)
summary(glmFR.water)
summary(glmVR.water)
summary(glmFV.water)
summary(glmRS.water)
summary(glmFS.water)
summary(glmVS.water)
summary(glmR.water)
summary(glmF.water)
summary(glmV.water)
summary(glmS.water)

quasi.water2$terms


summary(step(glm.water))
```
*This method is of comparing AIC values is too long and not worth it.*


## Using Anova function instead of comparing AICvalues
Since the above function did not work, and I needed to decide a model now, I tried comparing models that I thought would perform better by running the anova function while comparing their Mallow's Cp value.
```{r}
quasi.water2 <- glm(num_violations ~ num_facilities + num_site_visits + source + region, data = water, family = quasipoisson(link = "log"))

quasi.water2 %>%
  augment() -> quasi.resid
  
summary(quasi.water2)

quasi.water3 <- glm(num_violations ~ num_facilities + num_site_visits + source, data = water, family = quasipoisson(link = "log"))

summary(quasi.water3)
anova(quasi.water2, quasi.water3)

quasi.water4 <- glm(num_violations ~ num_facilities + source + region, data = water, family = quasipoisson(link = "log"))

summary(quasi.water4)

quasi.water5 <- glm(num_violations ~ num_facilities + num_site_visits +  region, data = water, family = quasipoisson(link = "log"))

quasi.null <- glm(num_violations ~ 1, data = water, family = quasipoisson(link = "log"))

anova(quasi.water, quasi.water2, test = "Cp")

anova(quasi.water, quasi.water2, test = "LRT")
```

After dropping the population parameter, a comparison of the Mallow's Cp and a Likelihood Ratio Test(LRT) ws done between the full and reduce model. Both show that the reduced model is better since the reduced model has a lower Mallow's Cp value and the p-value of the LRT test is 0.9185. This means we fail to reject the null hypothesis and we are able to say that population can be dropped.
